Dust trapping in thin-ringed protoplanetary disks

### 2018-07-27 projects/dullemond/paper_dusttraps/ms.tex
Added updated draft of dust trapping paper (still lots to do).

\abstract{A large fraction of the protoplanetary disks observed with the ALMA
  Large Programme {\bf XXXX} display multiple well-defined and nearly perfectly
  circular rings in the continuum, in many cases with substantial peak-to-valley
  contrast. Several of these rings are very narrow in radial extent.
  In this paper we analyze these results using the assumption that
  these dust rings are caused by dust trapping in radial pressure bumps. We
  model this process in a 1-D axisymmetric way, initially with a simple analytic
  model, then with a more detailed numerical model. We find that .....
%  including time-dependent viscous disk evolution coupled to time-dependent
%  radial dust drift. We perform model calculations without and with dust growth,
%  the latter with a simplified dust growth recipe. We test several scenarios
%  for the origin of the pressure bumps, including planetary gaps and dead zones.
{\bf CONCLUSIONS}
}

@@

The data from the ALMA Large Programme 2016.1.00484.L offers an exciting new
opportunity to put this concept to the test, and to put constraints on the
physics of dust trapping in axisymmetric pressure maxima. This is an opportunity
which we explore in this paper.

@@

\section{The high-contrast rings of AS 209, Elias 24, HD 163296, GW Lup and HD 143006}
\label{sec-data}
%
In this paper we focus on those sources of the ALMA Large Programme
that show high-contrast and radially thin rings that are separated by deep
valleys, and that are sufficiently face-on to not have to worry about 3-D
line-of-sight issues. These are AS 209, Elias 24, HD 163296, GW Lup and
HD 143006. Their stellar parameters are given in Table \ref{tab-stellar-params}.

\begin{table}
\begin{center}
\begin{tabular}{|c|ccc|}
\hline
\hline
Source     & $d\,\mathrm{[pc]}$ & $M_{*}/M_{\odot}$ & $L_{*}/L_{\odot}$ \\
\hline
AS 209     & 121              & 0.912        &  0.631 \\
Elias 24   & 136              & 0.776        &  6.03  \\
HD 163296  & 101              & 2.14         &  25.7  \\
GW Lup     & 155              & 0.457        &  0.331 \\
HD 143006  & 166              & 0.813        &  0.871 \\
\hline
\hline
\end{tabular}
\end{center}
\caption{\label{tab-stellar-params}The stellar parameters for the
  stars studied in this paper. Distance is in parsec and mass and luminosity
  are in units of the solar values.}
\end{table}

@@

The high-contrast rings of these sources provide ``clean laboratories''
for testing the theory of dust trapping, both in a ring-by-ring manner, as well
as by comparing to global dust drift models. Fig.\ \ref{fig-obs-profiles}
shows the radial profile (deprojected for
inclination) of the thermal emission of the dust of the five disks. The
procedure used to extract these radial profiles from the continuum maps
is described in {\bf [Paper Jane Huang, but note that for HD 163296 and
HD 143006 we removed the blob]}. The
brightness is expressed as linear brightness temperature
$T_{\mathrm{bright,linear}}=(c^2/2k_B)I_\nu/\nu^2$, where $I_\nu$ is the
intensity, $c$ the light speed, $k_B$ the Boltzmann constant, and $\nu$ the
frequency\footnote{It should be noted that the full (non-linear) brightness temperature,
defined such that $B_\nu(T_{\mathrm{bright,full}})=I_\nu$, with $B_\nu(T)$ the
Planck function, can have very different values, in particular at low
brightness. As a result $T_{\mathrm{bright,full}}(r)$ tends to vastly
underestimate the contrast between the rings and the inter-ring
regions. Therefore, in this paper we plot the brightness using the linear
brightness temperature, which is proportional to the intensity in Jy/beam.}.

@@

The eight rings which we study in this paper (see annotations in
Fig.~\ref{fig-obs-images}) are fitted with a Gaussian profile, shown as inverse
parabolas in Fig.\ \ref{fig-obs-profiles}. The fitting procedure was done in the
image plane, in order to make it possible to fit to a limited radial range,
i.e.\ selecting a single ring. Details of this fitting procedure, and tests
using the ALMA simulator, are described in Appendix {\bf XXXXX}. In
Fig.~\ref{fig-obs-gaussfits} these gaussian fits are shown in close-up. The
parameters of the best-fitting Gaussians are listed in Table
\ref{tab-gauss-params}.

@@

\begin{table*}
\begin{center}
\begin{tabular}{|cc|cc|ccccc|ccccc|}
\hline
\hline
%Source & Ring & $r_i\, [\mathrm{au}]$ & $r_o\, [\mathrm{au}]$ & $A\, [\mathrm{K}]$ & $r\, [\mathrm{au}]$ & $\mathrm{fwhm}\, [\mathrm{mas}]$ & $\sigma\, [\mathrm{au}]$ & $w_d\, [\mathrm{au}]$ & $T_{\mathrm{model}}\,\mathrm{[K]}$ & $w_d/h_p$ & $M_{\mathrm{d,ri}} [M_{\oplus}]$ & $M_{\mathrm{d,iv}} [M_{\oplus}]$\\
Source & Ring & $r_{\mathrm{in}}$ & $r_{\mathrm{out}}$ & $A$ & $r_0$ & $\mathrm{fwhm}$ & $\sigma$ & $w_d$ & $T_{\mathrm{model}}$ & $T_{\mathrm{model,br,lin}}$ & $w_d/h_p$ & $M_{\mathrm{d,ri}}$ & $M_{\mathrm{d,iv}}$\\
            & & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $[\mathrm{K}]$ & $[\mathrm{au}]$ & $[\mathrm{mas}]$ & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $\mathrm{[K]}$ & $\mathrm{[K]}$ &  & $[M_{\oplus}]$ & $[M_{\oplus}]$\\
\hline
% NOTE: This table is made automatically with python_make_tables/make_latex_table.py
AS 209     & 1 &  69 &  79 &   3.64 &  74.13 &  75.42 &   3.88 &   3.43 &  12.9 &   8.2 &   0.7 &  44.4 &   2.41\\
AS 209     & 2 & 115 & 125 &   2.95 & 120.31 &  90.28 &   4.64 &   4.28 &  10.1 &   5.6 &   0.5 & 101.9 &   5.54\\
Elias 24   & 1 &  72 &  82 &   5.22 &  76.85 &  87.99 &   5.08 &   4.66 &  22.3 &  17.2 &   0.6 &  41.1 &   2.23\\
HD 163296  & 1 &  60 &  76 &   8.48 &  67.41 & 171.73 &   7.36 &   7.21 &  34.2 &  29.0 &   1.6 &  50.4 &   2.74\\
HD 163296  & 2 &  94 & 104 &   4.92 &  99.86 & 133.56 &   5.73 &   5.53 &  28.1 &  22.9 &   0.8 &  42.5 &   2.31\\
GW Lup     & 1 &  79 &  89 &   1.48 &  85.13 &  72.08 &   4.74 &   4.15 &  10.3 &   5.7 &   0.6 &  36.3 &   1.97\\
HD 143006  & 1 &  35 &  45 &   3.21 &  41.05 &  71.76 &   5.06 &   4.42 &  18.8 &  13.8 &   1.7 &  16.7 &   0.91\\
HD 143006  & 2 &  59 &  72 &   2.48 &  64.83 & 111.83 &   7.88 &   7.49 &  15.0 &  10.1 &   1.7 &  43.5 &   2.36\\
\hline
\hline
\end{tabular}
\end{center}
\caption{\label{tab-gauss-params}The model parameters for the Gaussian ring fits
  in Figs.~\ref{fig-obs-profiles} and \ref{fig-obs-gaussfits}. The values $r_{\mathrm{in}}$
  and $r_{\mathrm{out}}$ are the inner and outer boundary of the fitting domain. $A$ is the
  peak brightness temperature of the ring, $r_0$ is its radius in $\mathrm{au}$,
  $\mathrm{fwhm}$ is the width expressed as full-width-at-half-maximum in units
  of milliarcsecond, $\sigma$ is its width expressed as standard deviation in
  units of $\mathrm{au}$ and $w_d$ is the width of the underlying (deconvolved)
  dust emission profile, also expressed as standard deviation in units of
  $\mathrm{au}$. The last five columns are derived from a simple disk model
  assumption: $T_{\mathrm{model}}$ is the midplane temperature computed from
  Eq.~(\ref{eq-disk-temperature-model}) assuming a flaring angle of
  $\varphi=0.02$. From that one can derive the corresponding linear brightness
  temerature $T_{\mathrm{model,br,lin}}=B_\nu(T_{\mathrm{model}})c^2/2\nu^2$ with
  $\nu=c/0.13\,\mathrm{cm}$ is the frequency belonging to the wavelength of
  0.13 $\mathrm{cm}$ of ALMA band 6.
  The third-to-last column expresses the deconvolved dust ring
  width in units of the disk pressure scale height $h_p$ computed from
  $T_{\mathrm{model}}$. The last two columns are the computed dust mass in the
  ring in units of Earth masses, assuming the small grain absorption opacity
  $\kappa_{\mathrm{abs,ri}}(\lambda=0.13\,\mathrm{cm})=1.70\,\mathrm{cm}^2/\mathrm{g}$
  of \citet{2010A&A...512A..15R} ($M_{\mathrm{d,ri}}$) and the corresponding
  value
  $\kappa_{\mathrm{abs,ri}}(\lambda=0.13\,\mathrm{cm})=31.3\,\mathrm{cm}^2/\mathrm{g}$
  of \citet{1997MNRAS.291..121I} ($M_{\mathrm{d,iv}}$). The stellar parameter
  are taken from Table \ref{tab-stellar-params}.}
\end{table*}

The observed rings are the result of the thermal emission of a dust ring convolved
with the ALMA beam. The standard deviation width $\sigma$ of the fitted Gaussian
profile is given in Table \ref{tab-gauss-params}.  To obtain the width of the
underlying dust ring we have to deconvolve. Assuming a Gaussian beam and a
Gaussian dust ring, we can use the rule of the convolution of two Gaussians, and
obtain the width $w_d$ of the dust ring
\begin{equation}
w_d=\sqrt{\sigma^2-b^2}
\end{equation}
where $b$ is the beam width expressed as standard deviation in units of
$\mathrm{au}$ ($b=d_{\mathrm{pc}}b_{\mathrm{fwhm,as}}/2.355$, where
$b_{\mathrm{fwhm,as}}$ is the beam full width at half maximum in units of
arcseconds, and $d_{\mathrm{pc}}$ is the distance to the source in units of
parsec). The width of the dust rings
in units of the local pressure scale height $h_p$ (last column of Table
\ref{tab-gauss-params}) can only be roughly estimated, because we do not know
the disk midplane temperature. Our estimate is made as follows:
\begin{equation}\label{eq-disk-temperature-model}
T_{\mathrm{model}}(r) = \left(\frac{\tfrac{1}{2}\varphi L_{*}}{4\pi r^2\sigma_{\mathrm{SB}}}\right)^{1/4}
\end{equation}
where $\sigma_{\mathrm{SB}}$ is the Stefan-Boltzmann constant and $\varphi$ is
the so-called flaring angle \citep[e.g.][]{1997ApJ...490..368C,
  1998ApJ...500..411D, 2001ApJ...560..957D}. The pressure scale height then
follows from $h_p=\sqrt{k_BT_{\mathrm{model}}r^3/\mu m_pGM_{*}}$ with $k_B$ the
Boltzmann constant, $m_p$ the proton mass, $G$ the gravitational constant and
$mu=2.3$ the mean molecular weight in atomic units. The stellar parameters are
taken from Table \ref{tab-stellar-params}. We take the flaring angle to be
$\varphi=0.02$ which is an estimate based on typical values from models.  But
this value can easily be wrong by a factor of four or so. The fact that the
pressure scale height goes as $T_{\mathrm{model}}^{1/2}\propto \varphi^{1/8}$
means, however, that such an error only affects $h_p$ mildly.

@@

\subsection{Ranges of $\alpha_{\mathrm{turb}}$, $\mathrm{Sc}$, $\mathrm{St}$ and $w$}
\label{sec-ranges-of-params}
%
First of all, the Schmidt number $\mathrm{Sc}$ is merely a way to relate the
turbulent viscosity with the turbulent mixing. If we do not strive to learn
about the turbulent viscosity and instead are satisfied with learning only about
the turbulent mixing, then we are only interested in the combination
$\alpha_{\mathrm{turb}}/\mathrm{Sc}$. For simplicity we set $\mathrm{Sc}=1$.

@@

The Stokes number $\mathrm{St}$ can be any value. But it is
directly related to the grain size $a$ and the gas density
$\rho_{\mathrm{g}}$. If we have observational constraints on $a$ and a good
estimate of the gas density $\rho_{\mathrm{g}}$, then we can eliminate this
uncertainty, and we are left with two unknown parameters
($\alpha_{\mathrm{turb}}$ and $w$) for one measurement ($w_{\mathrm{d}}$).
Unfortunately, while estimating $a$ from observations may be doable, it is far
more difficult to estimate $\rho_{\mathrm{g}}$. Standard disk gas mass estimates
are of limited use, as they are based on measuring the dust mass and multiplying
it by the estimated gas-to-dust ratio. Since we are testing the hypothesis of
dust trapping, we cannot assume a standard gas-to-dust ratio.

One can, however, set an upper bound on $\rho_{\mathrm{g}}$ by demanding that
the disk is gravitationally stable, i.e.\ that the Toomre parameter obeys
\begin{equation}
Q_{\mathrm{Toomre}}\equiv \frac{c_s\Omega_K}{\pi G \Sigma_{\mathrm{g}}} >2
\end{equation}
Here $c_s=\sqrt{k_BT/\mu m_p}$ is the isothermal sound speed, with $k_B$ the
Boltzmann constant, $m_p$ the proton mass, and $mu=2.3$ the mean molecular
weight in atomic units. $\Omega_K=\sqrt{GM_{*}/r^3}$ is the Kepler frequency,
$G$ is the gravitational constant and, finally, $\Sigma_{\mathrm{g}}$ the gas
surface density. The gas surface density and midplane density are related via
the pressure scale height $h_p=c_s/\Omega_K$ via the identity
$\Sigma_{\mathrm{g}}=\sqrt{2\pi}h_p\rho_{\mathrm{g}}$. Taking the disk midplane
temperature from Table \ref{tab-gauss-params}, which was calculated using
Eq.~(\ref{eq-disk-temperature-model}), we can compute the upper limits on
$\rho_{\mathrm{g}}$ for all of the rings. The results are given in Table
\ref{tab-ring-model-limits}.

>>>>

\abstract{A large fraction of the protoplanetary disks observed with the ALMA
  Large Programme DSHARP display multiple well-defined and nearly perfectly
  circular rings in the continuum, in many cases with substantial peak-to-valley
  contrast. Several of these rings are very narrow in radial extent.
  In this paper we analyze these results using the assumption that
  these dust rings are caused by dust trapping in radial pressure bumps. We
  model this process in a 1-D axisymmetric way, initially with a simple analytic
  model, then with a more detailed numerical model. We find that .....
%  including time-dependent viscous disk evolution coupled to time-dependent
%  radial dust drift. We perform model calculations without and with dust growth,
%  the latter with a simplified dust growth recipe. We test several scenarios
%  for the origin of the pressure bumps, including planetary gaps and dead zones.
{\bf CONCLUSIONS}
}

@@

The data from the ALMA Large Programme DSHARP ({\bf REFERENCES}) offers an exciting new
opportunity to put this concept to the test, and to put constraints on the
physics of dust trapping in axisymmetric pressure maxima. This is an opportunity
which we explore in this paper.

@@

\section{The high-contrast rings of AS 209, Elias 24, HD 163296, GW Lup and HD 143006}
\label{sec-data}
%
In this paper we focus on those sources of the DSHARP Programme
that show high-contrast and radially thin rings that are separated by deep
valleys, and that are sufficiently face-on to not have to worry about 3-D
line-of-sight issues. These are AS 209, Elias 24, HD 163296, GW Lup and
HD 143006. Their stellar parameters are given in Table \ref{tab-stellar-params}.

\begin{table}
\begin{center}
\begin{tabular}{|c|ccc|}
\hline
\hline
Source     & $d\,\mathrm{[pc]}$ & $M_{*}/M_{\odot}$ & $L_{*}/L_{\odot}$ \\
\hline
% AS 209     & 121              & 0.912        &  0.631 \\ % Data from 2018.05.13 Table
AS 209     & 121              & 0.832        &  1.413 \\ % 2018.07.02
Elias 24   & 136              & 0.776        &  6.03  \\ % 2018.07.02
% HD 163296  & 101              & 2.14         &  25.7  \\ % Data from 2018.05.13 Table
HD 163296  & 101              & 2.04         &  17.0  \\ % 2018.07.02
GW Lup     & 155              & 0.457        &  0.331 \\ % 2018.07.02
% HD 143006  & 166              & 0.813        &  0.871 \\ % Data estimated by Kees May/June 2018
HD 143006  & 165              & 1.78        &  3.80 \\  % 2018.07.02 ** Big change **
\hline
\hline
\end{tabular}
\end{center}
\caption{\label{tab-stellar-params}The stellar parameters assumed for the
  stars studied in this paper. Distance is in parsec and mass and luminosity
  are in units of the solar values.}
\end{table}

@@

The high-contrast rings of these sources provide ``clean laboratories''
for testing the theory of dust trapping in a ring-by-ring manner.
Fig.\ \ref{fig-obs-profiles}
shows the radial profile (deprojected for
inclination) of the thermal emission of the dust of the five disks. The
procedure used to extract these radial profiles from the continuum maps
is described in {\bf [Paper Jane Huang, but note that for HD 163296 and
HD 143006 we removed the blob]}. The
brightness is expressed as linear brightness temperature
$T_{\mathrm{br,lin}}=(c^2/2k_B)I_\nu/\nu^2$, where $I_\nu$ is the
intensity, $c$ the light speed, $k_B$ the Boltzmann constant, and $\nu$ the
frequency\footnote{It should be noted that the full (non-linear) brightness temperature,
defined such that $B_\nu(T_{\mathrm{br,full}})=I_\nu$, with $B_\nu(T)$ the
Planck function, can have very different values, in particular at low
brightness. As a result $T_{\mathrm{br,full}}(r)$ tends to vastly
underestimate the contrast between the rings and the inter-ring
regions. Therefore, in this paper we plot the brightness using the linear
brightness temperature, which is proportional to the intensity in Jy/beam.}.

@@

As is shown in Appendix \ref{sec-steady-state-analytic-trap-model}, for a
Gaussian pressure bump the solution to the radial dust mixing and drift problem
is, to first approximation, also a Gaussian, with a width smaller than, or equal
to, that of the gas pressure bump. Our analysis of the eight rings of this paper
therefore naturally starts with the fitting of the observed radial intensity
profiles with a Gaussian function. For some {\bf [CHECK]} of the rings of this
paper such Gaussian fits were done in the context of fitting the entire radial
intensity profile in the uv-plane to the ALMA data ({\bf Guzman et al.XXXX;
  More??}). Our fits may deviate slightly from those {\bf [CHECK]}, because here
we put emphasis on the shape of the peak of the rings, and deemphasize the
wings.

The fitting procedure is described in Appendix \ref{sec-gauss-fitting}.  The
fitting was done in the image plane, in order to make it possible to fit to a
limited radial range, i.e.\ selecting a single ring. The Gaussian fits are shown
as inverse parabolas in Fig.\ \ref{fig-obs-profiles}. In
Fig.~\ref{fig-obs-gaussfits} these Gaussian fits are shown in close-up. The
parameters of the best fits are listed in Table \ref{tab-gauss-params}.

@@

\begin{table*}
\begin{center}
\begin{tabular}{|cc|cc|ccccc|cccccc|}
\hline
\hline
%Source & Ring & $r_i\, [\mathrm{au}]$ & $r_o\, [\mathrm{au}]$ & $A\, [\mathrm{K}]$ & $r\, [\mathrm{au}]$ & $\mathrm{fwhm}\, [\mathrm{mas}]$ & $\sigma\, [\mathrm{au}]$ & $w_d\, [\mathrm{au}]$ & $T_{\mathrm{d}}\,\mathrm{[K]}$ & $w_d/h_p$ & $M_{\mathrm{d,ri}} [M_{\oplus}]$ & $M_{\mathrm{d,iv}} [M_{\oplus}]$\\
%Source & Ring & $r_{\mathrm{in}}$ & $r_{\mathrm{out}}$ & $A$ & $r_0$ & $\mathrm{fwhm}$ & $\sigma$ & $w_d$ & $T_{\mathrm{d}}$ & $T_{\mathrm{d,br,lin}}$ & $w_d/h_p$ & $M_{\mathrm{d,ri}}$ & $M_{\mathrm{d,iv}}$\\
%            & & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $[\mathrm{K}]$ & $[\mathrm{au}]$ & $[\mathrm{mas}]$ & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $\mathrm{[K]}$ & $\mathrm{[K]}$ &  & $[M_{\oplus}]$ & $[M_{\oplus}]$\\
Source & Ring & $r_{\mathrm{in}}$ & $r_{\mathrm{out}}$ & $A$ & $r_0$ & $\mathrm{fwhm}$ & $\sigma$ & $w_d$ & $T_{\mathrm{d}}$ & $T_{\mathrm{d,br,lin}}$ & $w_d/h_p$ & $\tau$ & $M_{\mathrm{d,ri}}$ & $M_{\mathrm{d,iv}}$\\
            & & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $[\mathrm{K}]$ & $[\mathrm{au}]$ & $[\mathrm{mas}]$ & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $\mathrm{[K]}$ & $\mathrm{[K]}$ &  &  & $[M_{\oplus}]$ & $[M_{\oplus}]$\\
\hline
% NOTE: This table is made automatically with python_make_tables/make_latex_table.py
% Note: New table made with new stellar parameters 2018.07.02
AS 209     & 1 &  69 &  79 &   3.64 &  74.13 &  75.42 &   3.88 &   3.43 &  15.8 &  10.9 &   0.6 &   0.41 &  33.3 &   1.81\\
AS 209     & 2 & 115 & 125 &   2.95 & 120.31 &  90.28 &   4.64 &   4.28 &  12.4 &   7.7 &   0.4 &   0.48 &  74.3 &   4.04\\
Elias 24   & 1 &  72 &  82 &   5.22 &  76.85 &  87.99 &   5.08 &   4.66 &  22.3 &  17.2 &   0.6 &   0.36 &  41.1 &   2.23\\
HD 163296  & 1 &  60 &  76 &   8.48 &  67.41 & 171.73 &   7.36 &   7.21 &  30.8 &  25.6 &   1.7 &   0.40 &  56.9 &   3.09\\
HD 163296  & 2 &  94 & 104 &   4.92 &  99.86 & 133.56 &   5.73 &   5.53 &  25.3 &  20.2 &   0.8 &   0.28 &  48.3 &   2.62\\
GW Lup     & 1 &  79 &  89 &   1.48 &  85.13 &  72.08 &   4.74 &   4.15 &  10.3 &   5.7 &   0.6 &   0.30 &  36.3 &   1.97\\
HD 143006  & 1 &  35 &  45 &   3.21 &  41.05 &  72.20 &   5.06 &   4.42 &  27.2 &  22.0 &   2.1 &   0.16 &  10.5 &   0.57\\
HD 143006  & 2 &  59 &  72 &   2.48 &  64.83 & 112.51 &   7.88 &   7.49 &  21.6 &  16.6 &   2.0 &   0.16 &  26.5 &   1.44\\
\hline
\hline
\end{tabular}
\end{center}
\caption{\label{tab-gauss-params}The model parameters for the Gaussian ring fits
  in Figs.~\ref{fig-obs-profiles} and \ref{fig-obs-gaussfits}. The values
  $r_{\mathrm{in}}$ and $r_{\mathrm{out}}$ are the inner and outer boundary of
  the fitting domain. $A$ is the peak linear brightness temperature of the
  best-fit Gaussian ring model, $r_0$ is its radius in $\mathrm{au}$,
  $\mathrm{fwhm}$ is the width expressed as full-width-at-half-maximum in units
  of milliarcsecond, $\sigma$ is its width expressed as standard deviation in
  units of $\mathrm{au}$ and $w_d$ is the width of the underlying (deconvolved)
  dust emission profile, also expressed as standard deviation in units of
  $\mathrm{au}$. The last six columns are derived from a simple disk model
  assumption: $T_{\mathrm{d}}$ is the midplane temperature of the disk (we
  assume gas and dust temperature to be equal) computed from
  Eq.~(\ref{eq-disk-temperature-model}) assuming a flaring angle of
  $\varphi=0.02$. From that one can derive the corresponding linear brightness
  temerature $T_{\mathrm{d,br,lin}}=B_\nu(T_{\mathrm{d}})c^2/2\nu^2$ where
  $\nu=c/0.13\,\mathrm{cm}$ is the frequency belonging to the wavelength of 0.13
  $\mathrm{cm}$ of ALMA band 6.  The fourth-to-last column expresses the
  deconvolved dust ring width in units of the disk pressure scale height $h_p$
  computed from $T_{\mathrm{d}}$. The $\tau$ column gives the estimated optical
  depth at the peak of the ring, calculated from
  $\tau=-\ln(1-A/T_{\mathrm{d,br,lin}})$.  The last two columns are the computed
  dust mass in the ring in units of Earth masses, assuming the small grain
  absorption opacity
  $\kappa_{\mathrm{abs,ri}}(\lambda=0.13\,\mathrm{cm})=1.70\,\mathrm{cm}^2/\mathrm{g}$
  of \citet{2010A&A...512A..15R} ($M_{\mathrm{d,ri}}$) and the corresponding
  value
  $\kappa_{\mathrm{abs,ri}}(\lambda=0.13\,\mathrm{cm})=31.3\,\mathrm{cm}^2/\mathrm{g}$
  of \citet{1997MNRAS.291..121I} ($M_{\mathrm{d,iv}}$). In making these
  estimates, the optically thin approximation was used. See Appendix
  \ref{sec-computing-dust-mass} for details. An estimate of the
  optical-depth-corrected dust mass can be made by multiplying the mass values
  in the table by $\tau/(1-e^{-\tau})$.}
\end{table*}

The observed rings are the result of the thermal emission of a dust ring convolved
with the ALMA beam. The standard deviation width $\sigma$ of the fitted Gaussian
profile is given in Table \ref{tab-gauss-params}.  To obtain the width of the
underlying dust ring we have to deconvolve. Assuming a Gaussian beam and a
Gaussian dust ring, we can use the rule of the convolution of two Gaussians, and
obtain the width $w_d$ of the dust ring
\begin{equation}
w_d=\sqrt{\sigma^2-b^2}
\end{equation}
where $b$ is the beam width expressed as standard deviation in units of
$\mathrm{au}$ ($b=d_{\mathrm{pc}}b_{\mathrm{fwhm,as}}/2.355$, where
$b_{\mathrm{fwhm,as}}$ is the beam full width at half maximum in units of
arcseconds, and $d_{\mathrm{pc}}$ is the distance to the source in units of
parsec). The width of the dust rings
in units of the local pressure scale height $h_p$ (see Table
\ref{tab-gauss-params}) can only be roughly estimated, because we do not know
the disk midplane temperature. Our estimate is made as follows:
\begin{equation}\label{eq-disk-temperature-model}
T_{\mathrm{d}}(r) = \left(\frac{\tfrac{1}{2}\varphi L_{*}}{4\pi r^2\sigma_{\mathrm{SB}}}\right)^{1/4}
\end{equation}
where $\sigma_{\mathrm{SB}}$ is the Stefan-Boltzmann constant and $\varphi$ is
the so-called flaring angle \citep[e.g.][]{1997ApJ...490..368C,
  1998ApJ...500..411D, 2001ApJ...560..957D}. The pressure scale height then
follows from $h_p=\sqrt{k_BT_{\mathrm{d}}r^3/\mu m_pGM_{*}}$ with $k_B$ the
Boltzmann constant, $m_p$ the proton mass, $G$ the gravitational constant and
$\mu=2.3$ the mean molecular weight in atomic units. The stellar parameters are
taken from Table \ref{tab-stellar-params}. We take the flaring angle to be
$\varphi=0.02$ which is an estimate based on typical values from models.  But
this value can easily be wrong by a factor of four or so. The fact that the
pressure scale height goes as $T_{\mathrm{d}}^{1/2}\propto \varphi^{1/8}$
means, however, that such an error only affects $h_p$ mildly.

@@

\subsection{Ranges of $\alpha_{\mathrm{turb}}$, $\mathrm{Sc}$, $\mathrm{St}$ and $w$}
\label{sec-ranges-of-params}
%
Reasonable values of $\alpha_{\mathrm{turb}}$, $\mathrm{Sc}$, $\mathrm{St}$ and
$w$ obey certain restrictions. First of all, the Schmidt number $\mathrm{Sc}$
is merely a way to relate the turbulent viscosity with the turbulent mixing. If
we do not strive to learn about the turbulent viscosity, and instead are
satisfied with learning only about the turbulent mixing, then we are only
interested in the combination $\alpha_{\mathrm{turb}}/\mathrm{Sc}$. For
simplicity we set $\mathrm{Sc}=1$.

@@

The Stokes number $\mathrm{St}$ can be any value. But it is directly related to
the grain size $a$ and the gas density $\rho_{\mathrm{g}}$, where the gas
density is directly related to the gas surface density $\Sigma_{\mathrm{g}}$ via
$\Sigma_{\mathrm{g}}=\sqrt{2\pi}h_p\rho_{\mathrm{g}}$. If we have observational
constraints on $a$ and a good estimate of the gas surface density
$\Sigma_{\mathrm{g}}$, then we can eliminate this uncertainty, and we are left
with two unknown parameters ($\alpha_{\mathrm{turb}}$ and $w$) for one
measurement ($w_{\mathrm{d}}$).  Unfortunately, while estimating $a$ from
observations may be doable, it is far more difficult to estimate
$\Sigma_{\mathrm{g}}$. Standard disk gas mass estimates are of limited use, as
they are based on measuring the dust mass and multiplying it by the estimated
gas-to-dust ratio. Since we are testing the hypothesis of dust trapping, we
cannot assume a standard gas-to-dust ratio.

One can, however, set an upper bound on $\Sigma_{\mathrm{g}}$ by demanding that
the disk is gravitationally stable, i.e.\ that the Toomre parameter obeys
\begin{equation}
Q_{\mathrm{Toomre}}\equiv \frac{c_s\Omega_K}{\pi G \Sigma_{\mathrm{g}}} >2
\end{equation}
Here $c_s=\sqrt{k_BT/\mu m_p}$ is the isothermal sound speed, with $k_B$ the
Boltzmann constant, $m_p$ the proton mass, and $mu=2.3$ the mean molecular
weight in atomic units. $\Omega_K=\sqrt{GM_{*}/r^3}$ is the Kepler frequency,
$G$ is the gravitational constant and, finally, $\Sigma_{\mathrm{g}}$ the gas
surface density. Taking the disk midplane
temperature from Table \ref{tab-gauss-params}, which was calculated using
Eq.~(\ref{eq-disk-temperature-model}), we can compute the upper limits on
$\Sigma_{\mathrm{g}}$ for all of the rings. The results are given in Table
\ref{tab-ring-model-limits}.
